---
title: "Assignment 3"
author: "Trishita Nath"
date: "2/13/2021"
output: html_document
---

## 1. Problem set 1
1. What is the rank of the matrix A?
```{r echo=TRUE, eval=TRUE}
#loading required libraries
library(pracma)
A <-matrix(c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3),nrow = 4,ncol = 4)
A
rref(A)
Rank(A)
```
2. Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?
The max rank cannot be > than ‘n’. The min rank in this case (non-zero) is 1.

Let Ra = rank of matrix X, where matrix X is an m x n matrix with m>n
Ra≤{min(row_rank,column_rank)Ra≤{min(m,n)}
We know the maximum rank will be n and since the matrix is non_zero,then
Ra≤nRa≥1
Hence minimum rank of matrix A would be 1
```{r eval=TRUE}
B <- matrix(c(1,3,2,2,6,4,1,3,2),nrow = 3, ncol= 3)
B
Rank(B)

```
## 2. Problem set 2
Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your
work. You’ll need to write out the characteristic polynomial and show your solution.
$$
A=\left[\begin{array}{cc} 
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6
\end{array}\right]\\
Ax=\lambda x\\
det(A-I\lambda)=0\\
$$
Solve for the roots of the polynomials to obtain the eigen values
$$
(1−\lambda)((4−\lambda)(6−\lambda))+2(0−0)+3(0−0)=0\\
(1−\lambda)(4−\lambda)(6−\lambda)−0+0=0\\
(1−\lambda)(24−4\lambda−6\lambda+λ^2)=0\\
(1−\lambda)(24−10\lambda+\lambda^2)=0\\
24−10\lambda+\lambda ^2−24λ+10\lambda ^2−\lambda ^3=0\\
−\lambda ^3+11\lambda ^2−34\lambda +24=0\\
$$
Divide every term on the left and right side of the equation by a -1 to make the leading term positive
$$
\lambda ^3-11\lambda ^2+34\lambda-24=0\\
$$
Factoring by grouping
$$
(\lambda^3-11\lambda^2)+(34\lambda-24)=0\\
\lambda ^ 2(\lambda-11)+2(17\lambda-12)=0\\
$$
This needs further factorization to solve. To simplify getting the roots of this equation, I will use R programming\
Verifying the characteristic polynomial:
```{r eval=TRUE}
A <- matrix(c(1,2,3,0,4,5,0,0,6), 3, 3, byrow=TRUE)
A
library(pracma)
charpoly(A)
```
Eigen values and vectors
```{r eval=TRUE}
eigen(A)
```